---
title: "Twitter Follower Mining"
output: html_notebook
---

This example pulls all of the followers for a given user and then does some text-mining on their descriptions:

* Generates a word cloud
* Breaks out their descriptions into "topics" using LDA (https://www.tidytextmining.com/topicmodeling.html)

This notebook uses `rtweet` for getting the Twitter data. The Shiny version of the app uses `twitteR` due to some limitations of `rtweet` in a Shiny environment.

## Setup/Config

```{r config}

# Set the base account
tw_account <- "garyvee"

# Load the necessary libraries. 
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rtweet,            # How we actually get the Twitter data
               tidyverse,         # Includes dplyr, ggplot2, and others; very key!
               knitr,             # Nicer looking tables
               tidytext,          # Tidy text!
               SnowballC,         # Mainly for stemming the search terms
               DT,                # Make a nice data table
               wordcloud,         # Word cloud creation
               RColorBrewer,      # Get some palettes to use with the word cloud
               topicmodels)       # For the topic modeling using LDA


# Minimum # of occurrences for a term to include in the wordcloud
min_frequency <- 2

# Set the number of topics to include in the topic model
num_topics <- 4

# Set the stopwords language
stopwords_lang <- "en"

# Words to exclude (because they're too dominant to be interesting). This will exclude
# them from the main word cloud AND from the topic modeling
exclude_words <- c("cancer", "breast", "de")

# Label for what this is
main_label <- paste0("Followers of @",tw_account)

# Get app credentials

# Name assigned to created app
tw_appname <- Sys.getenv("TWITTER_APPNAME")

# Key and Secret
tw_key <- Sys.getenv("TWITTER_KEY")
tw_secret <- Sys.getenv("TWITTER_SECRET")

# Create the token
tw_token <- create_token(
    app = tw_appname,
    consumer_key = tw_key,
    consumer_secret = tw_secret,
    set_renv = FALSE)

```

## Get the Data and Clean It Up

```{r get_data, message=FALSE, warning=FALSE}

# Get a list of all followers
user_followers <- get_followers(tw_account, n=10000, token = tw_token)

# # Split that vector up into a list with 15,000 users per (needed to get the details)
# user_followers_split <- split(user_followers, rep(1:ceiling(nrow(user_followers)/15000), each=15000, 
#                                                             length.out = nrow(user_followers)))

# Get the user details for all of those users
followers_details <-  lookup_users(user_followers$user_id, parse = TRUE, token = tw_token)

# Output how many users are being analyzed
cat(paste0("The following assessment covers the ", nrow(followers_details), " ", tolower(main_label), "."))

# Unnest it -- put each word on its own row and then collapse the individual
# words. This will also make everything lowercase and strip punctuation!
followers_data <- followers_details %>% 
  unnest_tokens(description_term, description) %>% 
  group_by(description_term) %>% 
  summarise(occurrences = n()) %>% 
  select(description_term, occurrences) %>% 
  ungroup() %>% 
  arrange(-occurrences)

# Remove the stop words. 1) get the stopwords, 2) remove 'em
stop_words <- get_stopwords(language = stopwords_lang) %>% 
  select(word)

followers_data <- followers_data %>% 
  anti_join(stop_words, by = c(description_term = "word"))

# Convert UTF-8 to ASCII (needed because all hell starts to break loose if you 
# try to text-mine multibyte). So, we're going to try to convert everything to
# ASCII. For some...this will fail and return NA. So, we'll then just remove
# the NA rows
followers_data <- followers_data %>%
  mutate(description_term = iconv(description_term, "UTF-8", "ASCII")) %>% 
  filter(!is.na(description_term))

# Perform stemming.
followers_data <- followers_data %>% 
  mutate(description_term_stem = wordStem(description_term))

# Go ahead and find the most popular un-stemmed word for each stemmed word.
# That will make the results look more "normal" to the casual viewer. We don't want
# to have any ties, so we're going to somewhat arbitrarily break any ties by adding
# the row number / 1000000 to each of the search counts first (We'll toss this later)
followers_data_top_term <- followers_data %>% 
  mutate(occurrences = occurrences + row_number()/1000000) %>% 
  group_by(description_term_stem) %>% 
  top_n(1, occurrences) %>% 
  select(-occurrences)

# Join that back to search data after totalling the occurrences by the stemmed term.
followers_data <- followers_data %>% 
  group_by(description_term_stem) %>% 
  summarise(occurrences = sum(occurrences)) %>% 
  left_join(followers_data_top_term) %>% 
  ungroup() %>% 
  select(description_term_stem, description_term, occurrences) %>% 
  arrange(-occurrences)

# Remove any additional "remove words" specified
followers_data <-  followers_data %>%
  filter(!description_term_stem %in% exclude_words)

# Get rid of the "top term" data frame
rm(followers_data_top_term)

```

## Show the Original Descriptions

```{r datatable_raw, message=FALSE, warning=FALSE}

followers_details %>% 
  arrange(-followers_count) %>% 
  select(user_id, description, followers_count) %>% 
  datatable(colnames = c("Username", "Description", "# of Followers"),  rownames = FALSE)

```


## Make a Term-Frequency Matrix

This looks similar to the report in Google Analytics, but it's been processed to be the individual words, stemmed, stopwords removed, etc.

```{r datatable_clean, message=FALSE, warning=FALSE}

select(followers_data, description_term, occurrences) %>% 
datatable(colnames = c("Description Term", "Occurrences"),
          rownames = FALSE)

```


## Create a Word Cloud

A wordcloud based on the cleaned up and unnested words.

```{r wordcloud, message=FALSE, warning=FALSE}

# Set a seed for reproducibility
set.seed(1971)

# Set a color palette
color_palette <- rev(brewer.pal(8,"Spectral")) 

# Generate the word cloud!
wordcloud(words = followers_data$description_term, 
          freq = followers_data$occurrences,
          scale=c(5.5,0.6),
          min.freq=min_frequency,
          max.words=500, 
          random.order=FALSE,
          rot.per=.0,
          colors=color_palette)

```

## Look for Topics!

We're going to use Latent Dirichlet allocation (LDA) to try to break out these words into topics. This is basically just following the process outlined for LDA at: https://www.tidytextmining.com/topicmodeling.html.

```{r lda, message=FALSE, warning=FALSE}

# Cast the term frequency matrix into a document term matrix. We're considering this all one 
# "document" so we're just hardcoding a "1" for that
followers_data_dtm <- followers_data %>% 
  mutate(doc = 1) %>% 
  cast_dtm(doc, description_term, occurrences)

# Run LDA. Setting a seed for reproducibility
search_lda <- LDA(followers_data_dtm, k = num_topics, control = list(seed = 1120))

# Assign a probability of each term being in each of the topics
search_topics <- tidy(search_lda, matrix = "beta")

# For each term, assign it to the topic for which it has the highest beta. This diverges
# from the approach described at tidytextmining.com, but it seems like a reasonably legit
# thing to do.
search_topics_and_terms <- search_topics %>%
  group_by(term) %>% 
  top_n(1, beta) %>% 
  ungroup() %>% 
  arrange(topic, -beta) %>% 
  left_join(followers_data, by = c(term = "description_term"))


# Function to generate a word cloud based on the topic ID passed in
generate_topic_wordcloud <- function(topic_id){
  
  # Filter the data to be just the topic and to 
  # knock out terms with a reallllly low beta
  topic_data <- search_topics_and_terms %>% 
    filter(topic == topic_id &
           beta > 0.001)
  
  # Generate the word cloud!
  wordcloud(words = topic_data$term, 
            freq = topic_data$occurrences,
            scale=c(3.5,1),
            min.freq=min_frequency,
            max.words=500, 
            random.order=FALSE,
            rot.per=.0,
            colors=color_palette)
}

# Call the function for each topic ID
topic_wordclouds <- map(seq(1:num_topics), generate_topic_wordcloud)


```

